---
title: "Unit 2 Homework"
author: "Jonathan A. Rocha"
date: "January 19, 2025"
output:
  word_document:
    fig_caption: true
    toc: true
    keep_md: true
  html_document: 
    toc: true
    df_print: tibble
    fig_caption: true
    number_sections: true
    keep_md: true
  pdf_document: 
    toc: true
    fig_caption: true
course: "DS 6371 - Statistics for Data Science"
professors: Dr. Bivin Sadler & Dr. Monnie McGee
editor_options:
  chunk_output_type: inline
---

# Homework Assignment: Unit 2

*Question 1:*

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)
library(ggplot2)
library(plotly)
library(tidyverse)
library(gridExtra)
library(moments)
library(nortest)
library(webshot2)

custom_dark_theme <- theme_dark() +
  theme(
    text = element_text(color = "#FFFFFF"),
    axis.text = element_text(color = "#FFFFFF"),
    axis.title = element_text(color = "#FFFFFF"),
    plot.title = element_text(color = "#FFFFFF"),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_line(color = "#333333"),
    plot.background = element_rect(fill = "#222222"),
    panel.background = element_rect(fill = "#222222")
  )

theme_set(custom_dark_theme)
```

## Introduction

The bumblebee bat (Kitti's hog-nosed bat) is recognized as the world's smallest mammal. This analysis examines weight measurements from a sample of these bats to test whether they come from a population with a mean weight of 1.8 grams.

## Data and Preliminary Analysis

`Input the bumblebee bat weight data`

```{r}
# Create a single data frame for consistency
bat_data <- data.frame(weight = c(1.7, 1.6, 1.5, 2.0, 2.3, 1.6, 1.6, 1.8, 1.5, 1.7, 1.2, 1.4, 1.6, 1.6, 1.6))
```

`Calculate comprehensive summary statistics`

```{r}
summary_stats <- data.frame(
  n = length(bat_data$weight),
  mean = mean(bat_data$weight),
  median = median(bat_data$weight),
  sd = sd(bat_data$weight),
  se = sd(bat_data$weight)/sqrt(length(bat_data$weight)),
  diff_from_null = mean(bat_data$weight) - 1.8,  # Difference from hypothesized mean
  skewness = moments::skewness(bat_data$weight),
  kurtosis = moments::kurtosis(bat_data$weight)
)

# Statistical tests
sw_test <- shapiro.test(bat_data$weight)
ad_test <- ad.test(bat_data$weight)
t_test_result <- t.test(bat_data$weight, mu = 1.8)
sign_test <- binom.test(
  sum(bat_data$weight > 1.8),
  length(bat_data$weight),
  p = 0.5,
  alternative = "two.sided"
)
```

`Print summary statistics with clear formatting`

```{r}
print("Summary Statistics:")
print(paste("Sample Size (n):", summary_stats$n))
print(paste("Sample Mean:", round(summary_stats$mean, 3), "grams"))
print(paste("Sample SD:", round(summary_stats$sd, 3), "grams"))
print(paste("Standard Error:", round(summary_stats$se, 3), "grams"))
print(paste("Difference from Null:", round(summary_stats$diff_from_null, 3), "grams"))
```

The sample mean of r round(summary_stats$mean, 3) grams is noticeably lower than our hypothesized value of 1.8 grams. This difference of r abs(round(summary_stats$diff_from_null, 3)) grams needs to be evaluated in the context of our sample's variability.

## Testing Assumptions

### Normality Assessment

`Create QQ plot`

```{r}
qq_plot <- ggplot(bat_data, aes(sample = weight)) +
  stat_qq(color = "#38bdf8") + 
  stat_qq_line(color = "#f87171") +
  labs(title = "Normal Q-Q Plot of Bumblebee Bat Weights",
       subtitle = paste("Shapiro-Wilk p-value:", round(sw_test$p.value, 4)),
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  annotate("text", x = -1, y = max(bat_data$weight),
           label = paste("Skewness:", round(summary_stats$skewness, 3)),
           color = "white", hjust = 0) +
  annotate("text", x = -1, y = max(bat_data$weight) - 0.1,
           label = paste("Kurtosis:", round(summary_stats$kurtosis, 3)),
           color = "white", hjust = 0) +
  custom_dark_theme +
  theme(plot.title = element_text(hjust = 0.5))

print(qq_plot)
```

The visual diagnostics suggest:

1.  The Q-Q plot shows points generally following the theoretical line, with some minor deviations at the extremes.
2.  The histogram reveals a roughly symmetric distribution with a slight right skew.
3.  We observe one potential outlier at 2.3g, but it's not extreme enough to warrant exclusion.

While not perfectly normal, the data appears to satisfy the normality assumption sufficiently for our t-test given:

-   The moderate sample size (n = 15)
-   The relatively symmetric distribution
-   The robustness of the t-test to minor departures from normality

`Create histogram with density curve`

```{r}
hist_plot <- ggplot(bat_data, aes(x = weight)) +
  geom_histogram(aes(y = after_stat(density)), binwidth = 0.1,
                fill = "#38bdf8", color = "black", alpha = 0.7) +
  geom_density(color = "#f87171", linewidth = 1) +
  geom_vline(xintercept = mean(bat_data$weight), 
             color = "#818cf8", linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = 1.8, 
             color = "#f87171", linetype = "dashed", linewidth = 1) +
  annotate("text", x = mean(bat_data$weight), y = max(density(bat_data$weight)$y),
           label = paste("x̄ =", round(mean(bat_data$weight), 3)),
           color = "white", hjust = -0.2) +
  annotate("text", x = 1.8, y = max(density(bat_data$weight)$y) * 0.9,
           label = "H₀: μ = 1.8",
           color = "#f87171", hjust = -0.2) +
  labs(title = "Distribution of Bumblebee Bat Weights",
       subtitle = paste("n =", length(bat_data$weight), ", SD =", round(sd(bat_data$weight), 3)),
       x = "Weight (grams)",
       y = "Density") +
  custom_dark_theme +
  theme(plot.title = element_text(hjust = 0.5))

print(hist_plot)
```

`Create t-distribution plot with critical regions`

```{r}
# Calculate t-statistic
t_stat <- (mean(bat_data$weight) - 1.8)/(sd(bat_data$weight)/sqrt(length(bat_data$weight)))
df <- length(bat_data$weight) - 1
t_crit <- qt(0.975, df)

t_vals <- seq(-4, 4, length.out = 1000)
t_dens <- dt(t_vals, df)
t_dist_data <- data.frame(t = t_vals, density = t_dens)

t_plot <- ggplot(t_dist_data, aes(x = t, y = density)) +
  geom_line(color = "#38bdf8") +
  geom_vline(xintercept = c(-t_crit, t_crit), 
             color = "#f87171", linetype = "dashed") +
  geom_vline(xintercept = t_stat, 
             color = "#818cf8", linetype = "solid") +
  geom_ribbon(data = subset(t_dist_data, t <= -t_crit | t >= t_crit),
              aes(ymax = density, ymin = 0), 
              fill = "#f87171", alpha = 0.2) +
  annotate("text", x = t_stat, y = max(t_dens) * 0.8,
           label = paste("t =", round(t_stat, 3)),
           color = "white") +
  annotate("text", x = -t_crit, y = max(t_dens) * 0.6,
           label = paste("t_crit = ±", round(t_crit, 3)),
           color = "#f87171") +
  labs(title = "t-Distribution with Critical Regions",
       subtitle = paste("p-value =", round(t_test_result$p.value, 4)),
       x = "t-value",
       y = "Density") +
 custom_dark_theme +
 theme(plot.title = element_text(hjust = 0.5))

print(t_plot)
```

`Create enhanced residual plot`

```{r}
residuals <- bat_data$weight - mean(bat_data$weight)
fitted_values <- rep(mean(bat_data$weight), length(bat_data$weight))

residual_plot <- ggplot(data.frame(fitted = fitted_values, residuals = residuals), 
                       aes(x = fitted, y = residuals)) +
  geom_point(color = "#38bdf8") +
  geom_hline(yintercept = 0, linetype = "dashed", color = "#f87171") +
  geom_smooth(method = "loess", color = "#818cf8", se = TRUE, fill = "#818cf8", alpha = 0.2) +
  annotate("text", x = mean(fitted_values), y = max(residuals),
           label = paste("SD of residuals:", round(sd(residuals), 3)),
           color = "white") +
  labs(title = "Residual Plot",
       subtitle = "Check for patterns",
       x = "Fitted Values",
       y = "Residuals") +
  custom_dark_theme +
  theme(plot.title = element_text(hjust = 0.5))

print(residual_plot)
```

`Arrange all plots in a grid`

```{r}
grid.arrange(qq_plot, hist_plot, t_plot, residual_plot,
            ncol = 2,
            top = grid::textGrob("Bumblebee Bat Weight Analysis", 
                               gp = grid::gpar(col = "white", fontsize = 16)))
```

`Create interactive QQ and histogram plots`

```{r warning=TRUE}
qq_interactive <- ggplotly(qq_plot) %>%
  layout(
    plot_bgcolor = "#222222",
    paper_bgcolor = "#222222",
    font = list(color = "#FFFFFF"),
    xaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Theoretical Quantiles",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    yaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Sample Quantiles",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    margin = list(l = 80, r = 40, t = 80, b = 80)  
  )

print(qq_interactive)

hist_interactive <- ggplotly(hist_plot) %>%
  layout(
    plot_bgcolor = "#222222",
    paper_bgcolor = "#222222",
    font = list(color = "#FFFFFF"),
    xaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Weight (grams)",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    yaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Density",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    margin = list(l = 80, r = 40, t = 80, b = 80)  
  )

print(hist_interactive)
```

`Print interactive plots side by side`

```{r}
plotly::subplot(qq_interactive, hist_interactive, nrows = 1, titleX = TRUE, titleY = TRUE, widths = c(0.5, 0.5)) %>%
  layout(
    plot_bgcolor = "#222222",
    paper_bgcolor = "#222222",
    font = list(color = "#FFFFFF"),
    margin = list(t = 50, b = 50, l = 50, r = 50),
    showlegend = FALSE,
    title = list(
      text = "Interactive Plots of Bumblebee Bat Weights",
      font = list(color = "#FFFFFF", size = 16)
    )
  )
```

## Step 1: State Hypotheses

The hypotheses can be written as:

The null hypothesis H₀ states that the population mean weight of bumblebee bats is 1.8 grams. The alternative hypothesis H₁ states that the population mean weight differs from 1.8 grams.

Mathematically, this can be expressed as:

-   H₀: μ = 1.8 (The population mean weight equals 1.8 grams)
-   H₁: μ ≠ 1.8 (The population mean weight differs from 1.8 grams)

## Step 2: Specify Significance Level

We will use α = 0.05 (two-tailed test) Critical values: t₀.₀₂₅,₁₄ = ±2.145 (from t-distribution with 14 df)

## Step 3: Calculate Test Statistic

`Calculate t-statistic`

```{r}
t_stat <- (mean(bat_data$weight) - 1.8)/(sd(bat_data$weight)/sqrt(length(bat_data$weight)))
```

`Print detailed t-statistic calculation`

```{r}
print(paste("Sample mean:", round(mean(bat_data$weight), 3)))
print(paste("Hypothesized mean:", 1.8))
print(paste("Standard error:", round(sd(bat_data$weight)/sqrt(length(bat_data$weight)), 4)))
print(paste("t-statistic:", round(t_stat, 4)))
```

## Step 4: Find P-value

`Calculate two-tailed p-value`

```{r}
p_value <- 2 * pt(abs(t_stat), df = length(bat_data$weight) - 1, lower.tail = FALSE)
print(paste("p-value:", round(p_value, 4)))

critical_value <- qt(0.975, df = length(bat_data$weight) - 1)
print(paste("Critical value (±):", round(critical_value, 4)))
```

`Print Comprehensive Statistical Summary`

```{r}
cat("\nComprehensive Statistical Analysis of Bumblebee Bat Weights\n")
cat("=======================================================\n\n")

cat("Basic Summary Statistics:\n")
cat("-----------------------\n")
cat("Sample Size:", summary_stats$n, "\n")
cat("Mean:", round(summary_stats$mean, 3), "grams\n")
cat("Median:", round(summary_stats$median, 3), "grams\n")
cat("Standard Deviation:", round(summary_stats$sd, 3), "grams\n")
cat("Standard Error:", round(summary_stats$se, 3), "grams\n")
cat("Skewness:", round(summary_stats$skewness, 3), "\n")
cat("Kurtosis:", round(summary_stats$kurtosis, 3), "\n\n")

cat("Normality Tests:\n")
cat("---------------\n")
cat("Shapiro-Wilk Test:\n")
cat("  W =", round(sw_test$statistic, 4), "\n")
cat("  p-value =", round(sw_test$p.value, 4), "\n\n")

cat("Anderson-Darling Test:\n")
cat("  A =", round(ad_test$statistic, 4), "\n")
cat("  p-value =", round(ad_test$p.value, 4), "\n\n")

cat("One-Sample t-Test Results:\n")
cat("------------------------\n")
cat("H₀: μ = 1.8 grams\n")
cat("t-statistic:", round(t_test_result$statistic, 3), "\n")
cat("degrees of freedom:", t_test_result$parameter, "\n")
cat("p-value:", round(t_test_result$p.value, 4), "\n")
cat("95% CI:", round(t_test_result$conf.int[1], 3), "to", 
    round(t_test_result$conf.int[2], 3), "grams\n\n")

cat("Sign Test Results:\n")
cat("----------------\n")
cat("H₀: median = 1.8 grams\n")
cat("Number of values > 1.8:", sign_test$statistic, "\n")
cat("p-value:", round(sign_test$p.value, 4), "\n")
```

## Step 5: Make Decision

With \|t-statistic\| = r round(abs(t_stat), 4) \> critical value = r round(critical_value, 4), and p-value = r round(p_value, 4) \< α = 0.05, we [reject the null hypothesis]{.underline}. There is sufficient evidence to conclude that the true population mean weight of bumblebee bats differs from 1.8 grams.

## Step 6: State Conclusion and Confidence Interval

`Calculate 95% confidence interval`

```{r}
t_crit <- qt(0.975, df = length(bat_data$weight) - 1)
margin_error <- t_crit * sd(bat_data$weight)/sqrt(length(bat_data$weight))
ci <- c(mean(bat_data$weight) - margin_error, mean(bat_data$weight) + margin_error)
print(paste("95% Confidence Interval:", round(ci[1], 3), "to", round(ci[2], 3)))
```

**Statistical Conclusion**: There is sufficient evidence to conclude that the true population mean weight of bumblebee bats differs from 1.8 grams (p = r round(p_value, 4)). The 95% confidence interval for the true population mean is (r round(ci[1], 3), r round(ci[2], 3)) grams.

**Practical Interpretation**: The sample data suggests that the hypothesized mean weight of 1.8 grams is plausible for this population. While our sample mean of r round(mean(bat_data\$weight), 3) grams differs slightly from 1.8, this difference is not statistically significant. The confidence interval tells us we can be 95% confident that the true population mean falls between r round(ci[1], 3) and r round(ci[2], 3) grams.

**Scope of Inference**: While our analysis provides valuable insights about bumblebee bat weights, the generalizability of these findings depends on the sampling methodology, which wasn't specified. If these measurements represent a random sample from the broader population, our conclusions could extend to the species as a whole. However, if the sample was taken from a specific location or time period, our inferences should be limited accordingly.
