---
title: "Statistical Analysis of Bumblebee Bat Weights"
author: "Jonathan A. Rocha"
date: "January 19, 2025"
output:
  word_document:
    fig_caption: true
    toc: true
    keep_md: true
  pdf_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
course: "DS 6371 - Statistics for Data Science"
professors: Dr. Bivin Sadler & Dr. Monnie McGee
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(plotly)
library(tidyverse)
library(gridExtra)

custom_dark_theme <- theme_dark() +
  theme(
    text = element_text(color = "#FFFFFF"),
    axis.text = element_text(color = "#FFFFFF"),
    axis.title = element_text(color = "#FFFFFF"),
    plot.title = element_text(color = "#FFFFFF"),
    panel.grid.major = element_line(color = "#444444"),
    panel.grid.minor = element_line(color = "#333333"),
    plot.background = element_rect(fill = "#222222"),
    panel.background = element_rect(fill = "#222222")
  )

theme_set(custom_dark_theme)
```

# Introduction

The bumblebee bat (Kitti's hog-nosed bat) is recognized as the world's smallest mammal. This analysis examines weight measurements from a sample of these bats to test whether they come from a population with a mean weight of 1.8 grams.

## Data and Preliminary Analysis

`Input the bumblebee bat weight data`

```{r}
bat_weights <- c(1.7, 1.6, 1.5, 2.0, 2.3, 1.6, 1.6, 1.8, 1.5, 1.7, 1.2, 1.4, 1.6, 1.6, 1.6)
```

`Calculate comprehensive summary statistics`

```{r}
summary_stats <- data.frame(
  n = length(bat_weights),
  mean = mean(bat_weights),
  sd = sd(bat_weights),
  se = sd(bat_weights)/sqrt(length(bat_weights)),
  diff_from_null = mean(bat_weights) - 1.8  # Difference from hypothesized mean
)
```

`Calculate basic summary statistics`

```{r}
bat_summary <- data.frame(
  n = length(bat_weights),
  mean = mean(bat_weights),
  sd = sd(bat_weights),
  se = sd(bat_weights)/sqrt(length(bat_weights)),
  t_crit = qt(0.975, df = length(bat_weights)-1)
)
```

`Print summary statistics with clear formatting`

```{r}
print("Summary Statistics:")
print(paste("Sample Size (n):", summary_stats$n))
print(paste("Sample Mean:", round(summary_stats$mean, 3), "grams"))
print(paste("Sample SD:", round(summary_stats$sd, 3), "grams"))
print(paste("Standard Error:", round(summary_stats$se, 3), "grams"))
print(paste("Difference from Null:", round(summary_stats$diff_from_null, 3), "grams"))
```

## Step 1: State Hypotheses

The hypotheses can be written as:

The null hypothesis H₀ states that the population mean weight of bumblebee bats is 1.8 grams. The alternative hypothesis H₁ states that the population mean weight differs from 1.8 grams.

Mathematically, this can be expressed as:

H₀: μ = 1.8 (The population mean weight equals 1.8 grams) H₁: μ ≠ 1.8 (The population mean weight differs from 1.8 grams)

and visualized with a histogram of the data:

`Create data frame for plotting`

```{r}
bat_df <- data.frame(weight = bat_weights)
```

## Testing Assumptions

### Normality Assessment

`Create QQ plot`

```{r}
ggplot(data.frame(bat_weights), aes(sample = bat_weights)) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Q-Q Plot of Bumblebee Bat Weights",
       x = "Theoretical Quantiles",
       y = "Sample Quantiles") +
  custom_dark_theme
```

`Create histogram with density curve`

```{r}
ggplot(data.frame(bat_weights), aes(x = bat_weights)) +
  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
  labs(title = "Distribution of Bumblebee Bat Weights",
       x = "Weight (grams)",
       y = "Frequency") +
  custom_dark_theme
```

`Convert to interactive plotly objects`

```{r}
qq_interactive <- ggplotly(qq_plot) %>%
  layout(
    plot_bgcolor = "#222222",
    paper_bgcolor = "#222222",
    font = list(color = "#FFFFFF"),
    xaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Theoretical Quantiles",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    yaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Sample Quantiles",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    margin = list(l = 80, r = 40, t = 80, b = 80)  
  )

hist_interactive <- ggplotly(hist_plot) %>%
  layout(
    plot_bgcolor = "#222222",
    paper_bgcolor = "#222222",
    font = list(color = "#FFFFFF"),
    xaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Weight (grams)",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    yaxis = list(
      gridcolor = "#444444",
      title = list(
        text = "Density",
        font = list(color = "#FFFFFF", size = 14)
      ),
      tickfont = list(color = "#FFFFFF")
    ),
    margin = list(l = 80, r = 40, t = 80, b = 80)  
  )
```

`Display plots side by side using subplot`

```{r}
subplot(qq_interactive, hist_interactive, nrows = 1, 
        titleX = TRUE, titleY = TRUE,  
        margin = 0.1) %>%  
  layout(
    showlegend = FALSE,
    title = list(
      text = "Bumblebee Bat Weight Distribution Analysis",
      font = list(color = "#FFFFFF", size = 16),
      y = 0.95
    ),
    margin = list(t = 100, b = 100),  
    width = 1200,   
    height = 600    
  )
```

## Step 2: Specify Significance Level

We will use α = 0.05 (two-tailed test) Critical values: t₀.₀₂₅,₁₄ = ±2.145 (from t-distribution with 14 df)

## Step 3: Calculate Test Statistic

`Calculate t-statistic`

```{r}
t_stat <- (mean(bat_weights) - 1.8)/(sd(bat_weights)/sqrt(length(bat_weights)))
```

`Print detailed t-statistic calculation`

```{r}
print(paste("Sample mean:", round(mean(bat_weights), 3)))
print(paste("Hypothesized mean:", 1.8))
print(paste("Standard error:", round(sd(bat_weights)/sqrt(length(bat_weights)), 4)))
print(paste("t-statistic:", round(t_stat, 4)))
```

## Step 4: Find P-value

`Calculate two-tailed p-value`

```{r}
p_value <- 2 * pt(abs(t_stat), df = length(bat_weights) - 1, lower.tail = FALSE)
print(paste("p-value:", round(p_value, 4)))
```

`Additional Context`

```{r}
critical_value <- qt(0.975, df = length(bat_weights) - 1)
print(paste("Critical value (±):", round(critical_value, 4)))
```

## Step 5: Make Decision

With \|t-statistic\| = r round(abs(t_stat), 4) \> critical value = r round(critical_value, 4), and p-value = r round(p_value, 4) \< α = 0.05, we [reject the null hypothesis]{.underline}.

## Step 6: State Conclusion and Confidence Interval

`Calculate 95% confidence interval`

```{r}
t_crit <- qt(0.975, df = length(bat_weights) - 1)
margin_error <- t_crit * sd(bat_weights)/sqrt(length(bat_weights))
ci <- c(mean(bat_weights) - margin_error, mean(bat_weights) + margin_error)
print(paste("95% Confidence Interval:", round(ci[1], 3), "to", round(ci[2], 3)))
```

**Statistical Conclusion**: There is sufficient evidence to conclude that the true population mean weight of bumblebee bats differs from 1.8 grams (p = r round(p_value, 4)). The 95% confidence interval for the true population mean is (r round(ci[1], 3), r round(ci[2], 3)) grams.

**Practical Interpretation**: The sample data suggests that the hypothesized mean weight of 1.8 grams is plausible for this population. While our sample mean of r round(mean(bat_weights), 3) grams differs slightly from 1.8, this difference is not statistically significant. The confidence interval tells us we can be 95% confident that the true population mean falls between r round(ci[1], 3) and r round(ci[2], 3) grams.

**Scope of Inference**: While our analysis provides valuable insights about bumblebee bat weights, the generalizability of these findings depends on the sampling methodology, which wasn't specified. If these measurements represent a random sample from the broader population, our conclusions could extend to the species as a whole. However, if the sample was taken from a specific location or time period, our inferences should be limited accordingly.
